Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-25 Paper/Paper.tex
fixed figure ordering
used small refs

What are the common features of pop music?  There are conventions
of subject matter, instrumentation, form, rhythm, harmony, and melody, among
others.  Our interest here is in the tonal content of the music -- \ie the
harmony and melody.  As a computationally-convenient proxy for a richer
description of the tonal content of audio, we use the popular Chroma
representation, which collapses an acoustic spectrum into a 12-dimensional
description, with one bin for each semitone of the western musical octave.
In addition, we simplify the time axis of our representation to take advantage
of the strong beat present in most pop music, and record just one
chroma vector per beat.  This beat-synchronous chromagram
representation represents a typical music track
in a few thousand values, yet when resynthesized back into audio via
modulation of octave-invariant ``Shepard tones'', the melody and
chord sequences of the original music usually remain recognizable
\cite{Ellis2007a}.  To the extent, then, that beat-chroma
representations preserve tonal content, they
are an interesting domain in which to search for patterns -- rich enough
to generate musically-relevant results, but simplified enough to
abstract away aspects of the original audio such as
instrumentation and other stylistic details.

@@

Prior work using small patches of chroma features includes
the ``shingles'' of \cite{Casey2007}, which were used to
identify ``remixes'', i.e., music based on some of the same
underlying instrument tracks, and also for matching
performances of Mazurkas \cite{Casey2008}.  That work,
however, was not concerned with extracting the deeper
common patterns underlying different pieces (and did not
use either beat- or bar-synchronous features).  Earlier work in
beat-synchronous analysis includes \cite{Bartsch2001},
which looked for repeated patterns within single songs to
identify the chorus, and \cite{Ellis2007a}, which 
cross-correlated beat-chroma matrices to match cover
versions of pop music tracks.  None of these works
examined or interpreted the content of the chroma
matrices in any detail.  In contrast, here we hope to
develop a codebook whose entries are of interest
in their own right.

@@

\subsection{Features}
The feature analysis used throughout this work is based on Echo Nest
analyze API \cite{EchoNest}.  
%
For any song uploaded to their platform this analysis returns a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars. Beats may span or 
subdivide segments; bars span multiple beats.
%
Averaging the per-segment chroma over beat times results in a
beat-synchronous chroma feature representation similar to that used in
\cite{Ellis2007a}.  Echo Nest chroma vectors are normalized to have the 
largest value in each column equal to 1.

>>>>

What are the common features of pop music?  There are conventions
of subject matter, instrumentation, form, rhythm, harmony, and melody, among
others.  Our interest here is in the tonal content of the music -- \ie the
harmony and melody.  As a computationally-convenient proxy for a richer
description of the tonal content of audio, we use the popular Chroma
representation, which collapses an acoustic spectrum into a 12-dimensional
description, with one bin for each semitone of the western musical octave.
In addition, we simplify the time axis of our representation to take advantage
of the strong beat present in most pop music, and record just one
chroma vector per beat.  This beat-synchronous chromagram
representation represents a typical music track
in a few thousand values, yet when resynthesized back into audio via
modulation of octave-invariant ``Shepard tones'', the melody and
chord sequences of the original music usually remain recognizable
\cite{Ellis2007a_small}.  To the extent, then, that beat-chroma
representations preserve tonal content, they
are an interesting domain in which to search for patterns -- rich enough
to generate musically-relevant results, but simplified enough to
abstract away aspects of the original audio such as
instrumentation and other stylistic details.

@@

Prior work using small patches of chroma features includes
the ``shingles'' of \cite{Casey2007_small}, which were used to
identify ``remixes'', i.e., music based on some of the same
underlying instrument tracks, and also for matching
performances of Mazurkas \cite{Casey2008}.  That work,
however, was not concerned with extracting the deeper
common patterns underlying different pieces (and did not
use either beat- or bar-synchronous features).  Earlier work in
beat-synchronous analysis includes \cite{Bartsch2001_small},
which looked for repeated patterns within single songs to
identify the chorus, and \cite{Ellis2007a_small}, which 
cross-correlated beat-chroma matrices to match cover
versions of pop music tracks.  None of these works
examined or interpreted the content of the chroma
matrices in any detail.  In contrast, here we hope to
develop a codebook whose entries are of interest
in their own right.

@@

\subsection{Features}
The feature analysis used throughout this work is based on Echo Nest
analyze API \cite{EchoNest}.  
%
For any song uploaded to their platform this analysis returns a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars. Beats may span or 
subdivide segments; bars span multiple beats.
%
Averaging the per-segment chroma over beat times results in a
beat-synchronous chroma feature representation similar to that used in
\cite{Ellis2007a_small}.  Echo Nest chroma vectors are normalized to have the 
largest value in each column equal to 1.
